Market Expectations and Mangerial Turnover

### 2024-04-08 paper_new.qmd
updated

```{r}
library(dplyr)
library(purrr)
#spells <- read.csv("raw_data/manager_spells_from_manager_urls_mgronly.csv")
spells <- read.csv("raw_data/manager_spells_from_manager_urls.csv")

# Define season_to_date() function here
season_to_date <- function(season) {
  if (is.na(season) | length(gsub("/", "", season)) == 0) {
    return(NA_real_)
  }

  parts <- strsplit(season, "/")[[1]]
  month <- as.numeric(parts[2])
   if (month < 8) {
     year <- paste0("20", substr(parts[1], 1, 2))
   } else {
     year <- paste0("19", substr(parts[1], 1, 2))
   }

  date_str <- paste0("01/08/", year)
  as.Date(paste(date_str, collapse = "/"), "%d/%m/%Y")
}

data_helper <- spells |>
  mutate(row_id=row_number()) |>
  filter(is.na(Finish)&contract_expiry!="-") |>
  select(contract_expiry, Finish,row_id)
data_helper$Finish<-as.Date(map_dbl(data_helper$contract_expiry, season_to_date))
spells |>
  mutate(row_id=row_number()) |>
  left_join(data_helper |> select(-Finish), by="row_id") |>
  mutate(Start = as.Date(Start),
         Finish = as.Date(Finish)) |>
  mutate(days_in_charge = as.numeric(Finish - Start)) ->spells
# find duplicates for name,club,Finish
#spells |> drop_na(Start,Finish) |> group_by(name,club,Finish) |> summarise(n=n()) |> filter(n>1) |> arrange(desc(n))
# remove duplicates
spells |> drop_na(Start,Finish) |> distinct(name,club,Finish,.keep_all = TRUE) |> filter(Finish-Start>0) -> spells

@@

```{r}
#|label: tbl-managertype
#| tbl-cap: "Manager types"
library(kableExtra)
terms_to_remove <- c("1.FC Köln II","Al-Ahli (UAE)","Gimnasia (J)","San Martín (T)",
  "San Martín (SJ)","Racing (Cba)","Juv. Unida (G)","1.FC Köln U19",
  "1.FC Köln U17","1.FC Köln Yth.",
  "1.FC Köln U16"," II","1.FC Köln","Bor. M'gladbach",
  "Hertha BSC","F. Düsseldorf",
  "Velez Mostar","SSV Reutlingen","FV Ravensburg",
  "Fenerbahce","FC Teningen","FV Nimburg","Heart of Midl.",
  "Vetra","Sp.Genzano (PZ)","Honvéd FC","Olympiacos","Honvéd SE","Hungary",
  "ADO Den Haag(A)","Dordrecht'90","Haarlem","AZ Alkmaar","Roda JC","AZ '67",
  "Sparta R.","De Treffers","NEC Nijmegen","FC Oss","VV Gemert","PSV U21")
spells %>%
  mutate(position = str_remove_all(position, paste0("(", paste(terms_to_remove, collapse = "|"), ")"))) -> spells
spells |> mutate(
    staff_position = trimws(tolower(position)),
    staff_position = ifelse(staff_position=="manager","manager",staff_position),
    staff_position = ifelse(staff_position=="assistant manager","assistant manager",staff_position),
    staff_position = ifelse(staff_position=="caretaker manager","caretaker manager",staff_position),
    staff_position = ifelse(str_detect(staff_position,"player-coach"),"player-coach",staff_position),
    staff_position = ifelse(str_detect(staff_position,"director of football"),"director of football",staff_position),
    staff_position = ifelse(str_detect(staff_position,"technical director"),"technical director",staff_position),
    staff_position = ifelse(str_detect(staff_position,"director of professional football"),"director of professional football",staff_position),
    staff_position = ifelse(str_detect(staff_position,"sporting director"),"sporting director",staff_position),
    staff_position = ifelse(str_detect(staff_position,"academy manager"),"academy manager",staff_position),
    staff_position = ifelse(str_detect(staff_position,"advisor"),"advisor",staff_position),
    staff_position = ifelse(str_detect(staff_position,"head of football operations"),"head of football operations",staff_position),
    staff_position = ifelse(str_detect(staff_position,"head of football operations"),"head of football operations",staff_position)) -> spells

spells %>%
  {table(.$staff_position[!is.na(.$staff_position)])} |> sort(decreasing = TRUE) |> 
  # make the frequency table into a data frame
  as.data.frame() |>
  # rename the columns
  rename("staff_position" = "Var1", "n" = "Freq") |>
  # order the data frame by the frequency
  arrange(desc(n)) |>
  head(20) |>
  kable(col.names=c("Staff Role","Number")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```

```{r}
#| label: fig-annual-spell-count
#| fig-cap: "Annual count of manager spell ending"
library(ggplot2)
library(dplyr)
# create plot of annual count of spells using a count of finishes per season assume season starts in August
spells |> 
  mutate(youth_team = ifelse(str_detect(club, "[A-Z][0-9]{2}|Yth."), "Yes", "No")) -> spells

# Create a list of the worlds countries in lowercase
countries <- c("albania", "andorra", "austria", "belarus", "belgium", "bosnia-herzegovina", "bulgaria", "croatia", "cyprus", "czech-republic", "denmark", "estonia", "faroe-islands", "finland", "france", "germany", "gibraltar", "greece", "hungary", "iceland", "ireland", "italy", "kosovo", "latvia", "liechtenstein", "lithuania", "luxembourg", "malta", "moldova", "montenegro", "netherlands", "north-macedonia", "northern-ireland", "norway", "poland", "portugal", "romania", "russia", "san-marino", "scotland", "serbia", "slovakia", "slovenia", "spain", "sweden", "switzerland", "turkey", "ukraine", "wales", "england")

# find the name of a country in the club name
spells |> 
  mutate(country = str_extract(club, "([A-Z][a-z]+)")) |> 
  filter(!is.na(country)) |> 
  count(country) |> 
  arrange(desc(n)) |> 
  head(20) |> 
  kable(col.names=c("Country","Number")) |> 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)

# plot the annual count of spells using a count of finishes per season assume season starts in August and condition on youth_team
spells |> 
  filter(staff_position=="manager") |> 
  mutate(season_start = ifelse(month(Start) < 8, year(Start), year(Start) + 1)) |>
  group_by(season_start,youth_team) |>
  count(season_start)

|> 
  ggplot(aes(x = season_start, y = n)) +
  geom_col(width = 0.5) +
  labs(title = "", x = "Season", y = "Number of spells") +
  theme_minimal() -> spell_count

plotly::ggplotly(spell_count)
```

```{python}
import pandas as pd
import time
from transformers import pipeline

generator = pipeline("text-generation")

# Load the unique club names from the CSV file
spells1 = pd.read_csv("raw_data/manager_spells_from_manager_urls_mgronly.csv")
club_names=spells1['club'].unique()

# Initialize an empty list to store the results
results = []

# Iterate over the array of club names and add a timer for each iteration
start_time = time.time()
for club in club_names:
    # Prepare the prompt for the model
    prompt = f"Provide information about the soccer club {club}: its country of origin,  division level, and whether it is a first  team or a youth team."

    # Generate the response using the model
    start = time.time()
    responses=generator(prompt)

    # Assuming the first (or only) response is what we're interested in
    response_text = responses[0]

    # Store the response, elapsed time for this iteration and total elapsed time in our results list
    results.append({'ClubName': club, 'Information': response_text, 'ElapsedTimePerIteration': time.time() - start, 
'TotalElapsedTime': time.time() - start_time})

# Convert the results list to a DataFrame
import pandas as pd
results_df = pd.DataFrame(results)

# Display the DataFrame to verify the results
print(results_df)

# Specify the file path for the CSV file
csv_file_path = 'soccer_clubs_information_hf.csv'

# Save the DataFrame to a CSV file
results_df.to_csv(csv_file_path, index=False)

print(f"Results have been saved to {csv_file_path}")

# Methodology

>>>>

```{r}
library(dplyr)
library(purrr)
#spells <- read.csv("raw_data/manager_spells_from_manager_urls_mgronly.csv")
spells <- readRDS("raw_data/manager_spells_from_manager_urls.rds")
names(spells)[20:22]<-c("Country","Division", "Full Name")
```

@@

```{r}
#|label: tbl-managertype
#| tbl-cap: "Manager types"
library(kableExtra)
spells %>%
  {table(.$staff_position[!is.na(.$staff_position)])} |> sort(decreasing = TRUE) |> 
  # make the frequency table into a data frame
  as.data.frame() |>
  # rename the columns
  rename("staff_position" = "Var1", "n" = "Freq") |>
  # order the data frame by the frequency
  arrange(desc(n)) |>
  head(20) |>
  kable(col.names=c("Staff Role","Number")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```

```{r}
#| label: fig-annual-spell-count
#| fig-cap: "Annual count of manager spell ending"
library(ggplot2)
library(dplyr)
# create plot of annual count of spells using a count of finishes per season assume season starts in August
spells |> 
  mutate(youth_team = ifelse(str_detect(club, "[A-Z][0-9]{2}|Yth.|Youth"), "Yes", "No")) -> spells

# create a list of european countries
european_countries <- c("albania", "andorra", "austria", "belarus", "belgium", "bosnia-herzegovina", "bulgaria", "croatia", "cyprus", "czech-republic", "denmark", "estonia", "faroe-islands", "finland", "france", "germany", "gibraltar", "greece", "hungary", "iceland", "ireland", "italy", "kosovo", "latvia", "liechtenstein", "lithuania", "luxembourg", "malta", "moldova", "montenegro", "netherlands", "north-macedonia", "northern-ireland", "norway", "poland", "portugal", "romania", "russia", "san-marino", "scotland", "serbia", "slovakia", "slovenia", "spain", "sweden", "switzerland", "turkey", "ukraine", "wales", "england")
# Create a list of north and south american countries
north_american_countries <- c("canada", "mexico", "united-states","USA")
# list of central american countries
central_american_countries <- c("belize", "costa-rica", "el-salvador", "guatemala", "honduras", "nicaragua", "panama")
south_american_countries <- c("argentina", "bolivia", "brazil", "chile", "colombia", "ecuador", "paraguay", "peru", "uruguay", "venezuela")
# Create list of australiasian countries
australasian_countries <- c("australia", "fiji", "new-zealand", "papua-new-guinea", "solomon-islands", "vanuatu")
# Create list of african countries
african_countries <- c("algeria", "angola", "benin", "botswana", "burkina-faso", "burundi", "cameroon", "cape-verde", "central-african-republic", "chad", "comoros", "congo", "djibouti", "egypt", "equatorial-guinea", "eritrea", "eswatini", "ethiopia", "gabon", "gambia", "ghana", "guinea", "guinea-bissau", "ivory-coast", "kenya", "lesotho", "liberia", "libya", "madagascar", "malawi", "mali", "mauritania", "mauritius", "morocco", "mozambique", "namibia", "niger", "nigeria", "rwanda", "sao-tome-and-principe", "senegal", "seychelles", "sierra-leone", "somalia", "south-africa", "south-sudan", "sudan", "tanzania", "togo", "tunisia", "uganda", "zambia", "zimbabwe")
# Create list of asian countries
asian_countries <- c("afghanistan", "armenia", "azerbaijan", "bahrain", "bangladesh", "bhutan", "brunei", "cambodia", "china", "east-timor", "georgia", "india", "indonesia", "iran")
# Create a list of countries that are not in the above lists
other_countries <- c("anguilla", "antigua-and-barbuda", "aruba", "bahamas", "barbados", "bermuda", "british-virgin-islands", "cayman-islands", "curacao", "dominica", "falkland-islands", "greenland", "grenada", "guadeloupe", "haiti", "jamaica", "martinique", "montserrat", "netherlands-antilles", "puerto-rico", "saint-kitts-and-nevis", "saint-lucia", "saint-vincent-and-the-grenadines", "sint-maarten", "trinidad-and-tobago", "turks-and-caicos-islands", "us-virgin-islands")
all_countries <- c(european_countries, north_american_countries, central_american_countries, south_american_countries, australasian_countries, african_countries, asian_countries, other_countries)
spells |> mutate(national_team = if_else(tolower(club) %in% all_countries,"Yes","No")) -> spells
# plot the annual count of spells using a count of finishes per season assume season starts in August and condition on youth_team
spells |> 
  filter(staff_position=="manager" & national_team =="No" & youth_team=="No") |> 
  mutate(season_start = ifelse(month(Start) < 8, year(Start), year(Start) + 1)) |>
  group_by(season_start,national_team) |>
  count(season_start)  |> 
  ggplot(aes(x = season_start, y = n, fill=national_team)) +
  geom_col(width = 0.5) +
  labs(title = "Manager-Club spells per annum", x = "Season Start Year", y = "Number of Turnovers") +
  theme_minimal() -> spell_count
plotly::ggplotly(spell_count)
```

```{r}
spells |> 
  filter(staff_position=="manager" & national_team =="No" & youth_team=="No") |> 
  mutate(season_start = ifelse(month(Start) < 8, year(Start), year(Start) + 1)) |>
  group_by(season_start,national_team) |>
  count(season_start)  |> 
  ggplot(aes(x = season_start, y = n, fill=national_team)) +
  geom_col(width = 0.5) +
  labs(title = "Manager-Club spells per annum", x = "Season Start Year", y = "Number of Turnovers") +
  theme_minimal() -> spell_count
plotly::ggplotly(spell_count)
```

```{r}
spells |>
```
# Methodology
